---
output: rmarkdown::github_document
bibliography: README-refs.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# R/`tmle3mopttx`

[![Project Status: WIP - Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](http://www.repostatus.org/badges/latest/wip.svg)](http://www.repostatus.org/#wip)

> Targeted Learning and Variable Importance with Optimal Individualized Categorical Treatment

__Authors:__ [Ivana Malenica](https://github.com/podTockom), [Jeremy R. Coyle](https://github.com/jeremyrcoyle) and [Mark J. van der Laan](https://vanderlaan-lab.org/)

## Description

`tmle3mopttx` is an adapter/extension R package in the `tlverse` ecosystem, that supports estimation of the optimal individualized treatment and the mean outcome under the estimated rule. In particular, it relies on the modern implementation of the Super Learner algorithm, [sl3](https://github.com/tlverse/sl3) and [tmle3](https://github.com/tlverse/tmle3) for the targeting step [@vdl2007super, @vdl2011targeted]. 

In order to avoid nesting cross-validation, it relies on split-specific estimates of $\mathbb{E}(Y|A,W)$ and $P(A|W)$ in order to estimate the rule, as described by Coyle et al [@jeremythesis]. In addition, the targeted maximum likelihood estimates of the mean performance under the estimated rule are obtained using CV-TMLE [@cvtmle2010]. The implementation supports categorical treatments, providing three different versions for the rule estimation. The goal of this work is to build upon the tlverse framework and the estimation methodology implemented for a single mean counterfactual outcome in order to introduce an end-to-end methodology for variable importance analyses.

---

## Installation

You can install the most recent _stable release_ from GitHub via [`devtools`](https://www.rstudio.com/products/rpackages/devtools/) with:
                                                                                              
```{r tmle3mopttx-installation, eval = FALSE}
devtools::install_github("tlverse/tmle3mopttx")
```

---

## Issues

If you encounter any bugs or have any specific feature requests, please [file an
issue](https://github.com/tlverse/tmle3mopttx/issues).

---

## References

