---
output:
  rmarkdown::github_document
bibliography: "README-refs.bib"
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# R/`tmle3mopttx`

[![Travis-CI Build Status]()
[![AppVeyor Build  Status]()
[![Coverage Status]()
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](http://www.repostatus.org/badges/latest/wip.svg)](http://www.repostatus.org/#wip)
[![License: GPL v3](https://img.shields.io/badge/License-GPL%20v3-blue.svg)](http://www.gnu.org/licenses/gpl-3.0)

> Targeted Learning and Variable Importance with Optimal Individualized Categorical Treatment

__Authors:__ [Ivana Malenica](https://github.com/podTockom), [Jeremy R. Coyle](https://github.com/jeremyrcoyle) and [Mark J. van der Laan](https://vanderlaan-lab.org/)

## Description

`tmle3mopttx` estimates the optimal individualized treatment rule for the categorical treatment using the modern implementation of the Super Learner algorithm ([sl3](https://github.com/tlverse/sl3)) and [tmle3](https://github.com/tlverse/tmle3) [@vdl2007super]. In order to avoid nesting cross-validation, it relies on split-specific estimates of $Q$ ($\mathbb{E}(Y|A,W)$) and $g$ (($P(A|W)$)) in order to estimate the rule as described by Coyle et al [@jeremythesis]. In addition, it provides the Targeted Maximum Likelihood estimates of the mean performance using CV-TMLE under the estimated rules. This is an adapter package for use with the `tmle3` framework and the tlverse software ecosystem for Targeted Learning.

---

## Installation

You can install the most recent _stable release_ from GitHub via
[`devtools`](https://www.rstudio.com/products/rpackages/devtools/) with:

```{r gh-installation, eval = FALSE}
devtools::install_github("tlverse/tmle3mopttx")
```

---

## Issues

If you encounter any bugs or have any specific feature requests, please [file an
issue](https://github.com/tlverse/tmle3mopttx/issues).

---

## License


The contents of this repository are distributed under the GPL-3 license. See
file `LICENSE` for details.
